Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2012-12-04 doc/paper-lcr.tex
updated figure2
response to Tom's edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/cornerplot.eps}
\caption[Biased?]{The conditional distribution of 
$\kappax$ given $\kappah$, in the case where  
we have perfect knowledge of halo mass and redshift.
$10^5$ reconstructed lines of
sight were used to make this plot. 
Solid (dashed) lines enclose 68\% (95\%) of the conditional probability.
$\kappah$ traces $\kappax$, but with a positive offset which arises from
$\kappah$ not accounting for any voids or dark structures.
At fixed $\kappah$ the scatter in $\kappax$ grows rapidly with $\kappah$.
\flag{Can you plot a $y=x$ line to show the correction being made, please?}}
\label{fig:jointkh-k}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\tom{With photometric data alone, $\kappah^{\rm med}$ can shift by $\sim$0.01 relative to $\kappah^{\rm med}$
given spectroscopic coverage. The $\kappax$ contribution of voids cannot change with data quality;
the shift in $\kappah^{\rm med}$ is due to the asymmetric propogation of stellar mass uncertainties
into $\kappax$ uncertainties. As stellar mass uncertainty increases $\kappah^{\rm med}$ is pushed higher. 
To correctly calibrate  $\pr(\kappax|\kappah^{\rm med},\data)$ one {\it must} include the data quality used to generate $\kappah^{\rm med}$.}

@@

\comment{From Sherry:
Re: $P_N$:
Not sure that this is the best way to quantify bias.  First of all, this
doesn't include the modeling error of $\sim3-4\%$, so the number of lenses listed
(for when <$P_N$> = half width of $P_N$) is underestimated.  Also, with this
definition, a large number of lenses could be described as "biased" even if
the resulting shift is say 0.0001, which won't really trouble us at all.  For
example, what would the distributions look like in Fig.8 with 100 lenses? 
Since the peaks of the red curves are within 0.01 of kappa_true, I would
suspect that with a large number of lenses, the corresponding red curve would
be within 0.01 of kappa_true.  The present day goal is to get H0 to 1\%, and
it would be good to state somewhere that with targeted spectroscopy, one could
recover kext at the 1\% level with a large sample of lenses.}

@@

We first consider a purely random selection of lines of sight. With the purely
photometric reconstruction, there is no evidence of bias, even after combining
20 lines of sight. \tom{This result is a sanity check: if we had chosen lines of
sight with identical $\kappah^{\rm med}$ rather than $\kappah^{\rm med}\pm0.003$ randomly
selected lines of sight would have zero bias by construction}. If lines of sight hosting time delay lenses
are truely random,  our method allows $\kappax$ to be reconstructed without inducing a bias\footnote{under the assumption that the universe is like the calibration simulation, and the correct stellar-mass to halo-mass relation has been used}.

@@

We find that selecting the most tightly constrained lines of sight in this way
does not introduce a bias: 
our reconstruction is most accurate for the lowest
$\kappax$ lines of sight. Combining 20 such systems the bias is at the 0.001 level.

@@

\phil{We can test the impact of such a shear selection by selecting only lines
of sight with an external shear above some threshold.  We define a somewhat
extreme selection  of $\gammax>0.05$, and find that in this sample, $\kappax$
would be systematically underestimated at the $\sim$0.008 level (corresponding to a
0.8\% systematic error in distance).} 
In the absence of other sources of uncertainty, this
systematic error would be significant in a sample of just 5 lenses; 
approximating strong-lens
modelling uncertainties as an unbiased Gaussian with width 4\% on time-delay distance a 0.8\% bias would be
significant for $\sim$20 lenses.
This systematic error is mitigated if only
the lines of sight with both $\gammax>0.05$ and well constrained $\pr(\kappax|\data)$ are
used, but in this case a bias at the $\sim$0.005 level remains.

@@

\comments{
We have shown that reconstructing the matter due to halos along any line of
sight can improve the constraints on the external convergence along that line
of sight, with, depending on sample selection,  the width of
$\pr(\kappax|\data)$ being up to a factor of two smaller than the global
$\pr(\kappax)$, and  we have quantified the residual bias in the $\kappax$
estimates under our calibration assumptions. We now revisit these assumptions.
}

@@

Filaments and voids are more difficult to model, but simple forms could in
principle be derived from the structures found in the simulations, and
included in the halo model in some way. Most importantly for the highest
$\kappax$ lines of sight will be a better group and cluster model:
\citet{MomchevaEtal2006} and  \citet{WongEtal2011} recommend using models
where dark mass is assigned to both galaxy-scale halos and halos for the
groups and clusters they occupy: to make this work, one could incorporate a
group and cluster finder, and then infer halo mass from optical richness
\citep[\eg][]{MaxBCG} in tandem with BCG stellar mass.

\comments{Interestingly, we have found that it is the most empty lines of sight that 
can be reconstructed with the highest precision. $\kappax$ estimates for empty
lines of sight receive few contributions from halos, and a large contribution
from voids; since they have the tightest PDFs after the reconstruction it
seems that our model's biggest uncertainties are driven by naively
reconstructing halos rather than neglecting voids. With a photometric
reconstruction of the field there is a significant broadening, but the
resulting precision is still typically 50\%  higher than using the global
$\kappax$ distribution.

Since the uncertainty on $\pr(\kappax|\data)$ is $\sim$0.01 even
with perfect knowledge of halo mass and redshift, it seems that modelling uncertainties
from voids, deviations from sphericity and dark matter clumping within the
main halo dominate the total $\kappax$ uncertainty even with very
uncertain stellar masses. Inferring halo ellipticity and dark matter
clumping will likely always remain a difficulty for line of sight
reconstruction; but as \Fref{fig:magcut} shows, observing deeper
than 24th magnitude does not significantly help the reconstruction.
}

\comments{For a small fraction of lines of sight, $\pr(\kappax|\data)$ remains very
broad even after applying our reconstruction; these lines of sight are
typically the most over-dense lines of sight in the universe. For
time-delay cosmography the most observationally expensive tasks are the
light-curve monitoring and high resolution follow-up imaging and spectroscopy,
while making photometric observations of a
4$\times$4 arcminute patch of sky survey down to 24th magnitude is
relatively cheap. We have shown that with a single epoch observation of
the region around a strong lens it is possible to infer $\pr(\kappax|\data)$.
If a line of sight has a broad $\pr(\kappax|\data)$ it can be rejected {\it
before} the investment of follow-up. We find that
selecting only the lines of sight with the most tightly constrained
$\pr(\kappax)$ does not induce a bias in $\kappax$, and hence distance.
For the most over-dense lines of sight,
the precision of $\pr(\kappax|\data)$ is improved noticeably by the addition
of spectroscopic information; currently insufficient time-delay systems are
know to choose a subset based on their reconstructed $\pr(\kappax|\data)$.
Spectroscopic coverage will definitely be useful for extracting cosmology from
over-dense lines of sight.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\comments{%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/biasplots3.eps}
\caption{Convergence estimation accuracy using the full  $\pr(\kappax|\kappah)$
distribution, rather than its median.  As in \Fref{fig:biasplots}, solid,
dashed and dotted lines show combinations of 20, 5 and 2 lines of sight
respectively, while green lines show the inferences 
from photometry alone, red lines from photometry and targeted spectroscopy 
(assuming a spectroscopic redshift for all
halos with $i<23$ and all halos within 1 arcminute and $i<24$), and blue lines
from the unrealistic assumption of a spectroscopic redshift for all galaxies
in the field. }
\label{fig:no-median-tricks}
\end{figure}
}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\comments{ One
significant disadvantage of these types of method is that they impose a
bottleneck in the flow of information. Once you have compressed to one number,
it is difficult to incorporate more information later; you have to recompute
all the calibrations. Using the full PDF for $\kappah$ is the more correct
thing to do: at present, however, it suffers from significant bias in the
photometric data case. \Fref{fig:no-median-tricks} shows how $\kappax$ is
overestimated, as a result of the high scatter in halo mass for objects with 
high stellar mass. As has already been noted, 
future work should focus on increasing the precision of the halo
mass estimates. 
}

In this work, we have used  the \MS as a calibration tool. If the real
Universe is not like the \MS, then this calibration will be incorrect.
Repeating this  study using a different simulation to make the test catalogues
would allow us to quantify the sensitivity of our results to the simulation
used. Similarly the use of a different semi-analytic model to paint galaxies
onto halos, and an alternative stellar mass--halo mass relation would both
enable similar exploration of these systematic errors. \citet{SuyuEtal2010}
used the relative over-density to mitigate this effect, but it does not
completely remove the dependence on simulation.

\tom{ Future work should focus on two areas:
reducing the dependance on simulations; and,
improving the modeling of dark structures and voids. The analysis should be repeated with different 
simulations and differrent stellar-mass to halo-mass relations, cross-checking that 
using the wrong simulation simulation or relation does not induce a systematic error.
Improving the modelling of of voids could be done using the void models of \citet{voids2}
or \citet{voids1} combined with a void finder such as in \citet{voids3}.
Combining the halo model with a void model may allow a direct test of the 
halo approximation; using real photometric data one could reconstruct lines of sight
and compare against measured weak lensing shears. This would ensure that the halo model reproduces the
contents of the real universe, rather than a particular simulation.}

\comments{In particular, the method presented here is tied to the simulations at two
points: the prior PDF for $\kappah$, and the conditional distribution
$\pr(\kappax|\kappah)$ used to convert halo model convergence to total
convergence. The first tie could potentially be avoided if the halo mass
estimation gave rise to a reasonable $\pr(\kappah|\data)$ on its own.
Possibilities for improving the precision of the halo mass estimation are
discussed above. The second tie is more difficult, since breaking it means
modelling the unseen mass structures and voids. It
should be possible to infer hierarchically  the parameters of a flexible model
for these things from the data itself, as demonstrated by
\citet{KarpenkaEtal2012} (albeit under the assumption of scatter-free scaling
relations) -- how this affects the cosmographic precision is to be
investigated. When working with wide field imaging surveys,  weak lensing shears will be
available around every line of sight:  fitting these with the halo models
described here in order to constrain the hyper-parameters of the model presents a 
possible test of our halo model prescription against real observations.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\item Selecting from the highest precision 33\% of the reconstructed lines of
sight results in a sample with low bias, $\sim$0.1\% in terms of the time-delay
distance. \tom{For these lines of sight the reconstructed $\pr(\kappax)$ induce 
a statistical uncertainty of $\sim$1\% on individual time-delay distances}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/cornerplot.eps}
\caption[Biased?]{The conditional distribution of 
$\kappax$ given $\kappah$, in the case where  
we have perfect knowledge of halo mass and redshift.
$10^5$ reconstructed lines of
sight were used to make this plot. 
Solid (dashed) lines enclose 68\% (95\%) of the conditional probability.
$\kappah$ traces $\kappax$, but with a positive offset which arises from
$\kappah$ not accounting for any voids or dark structures.
At fixed $\kappah$ the scatter in $\kappax$ grows rapidly with $\kappah$.
\tom{The thin diagonal line follows $\kappax=\kappah$.}}
\label{fig:jointkh-k}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\tom{With photometric data alone, $\kappah^{\rm med}$ can shift by $\sim$0.01
relative to $\kappah^{\rm med}$ given spectroscopic coverage. Since  the
$\kappax$ contribution of voids cannot change with data quality, the shift in
$\kappah^{\rm med}$ must be due to the asymmetric propagation of stellar mass
uncertainties into $\kappax$ uncertainties. As the stellar mass uncertainty
increases, $\kappah^{\rm med}$ is pushed higher.  To correctly calibrate 
$\pr(\kappax|\kappah^{\rm med},\data)$ one {\it must} include the data quality
used to generate $\kappah^{\rm med}$.}

@@

% \comment{From Sherry:
% Re: $P_N$:
% Not sure that this is the best way to quantify bias.  First of all, this
% doesn't include the modeling error of $\sim3-4\%$, so the number of lenses listed
% (for when <$P_N$> = half width of $P_N$) is underestimated.  Also, with this
% definition, a large number of lenses could be described as "biased" even if
% the resulting shift is say 0.0001, which won't really trouble us at all.  For
% example, what would the distributions look like in Fig.8 with 100 lenses? 
% Since the peaks of the red curves are within 0.01 of kappa_true, I would
% suspect that with a large number of lenses, the corresponding red curve would
% be within 0.01 of kappa_true.  The present day goal is to get H0 to 1\%, and
% it would be good to state somewhere that with targeted spectroscopy, one could
% recover kext at the 1\% level with a large sample of lenses.}

@@

We first consider a purely random selection of lines of sight. With the purely
photometric reconstruction, there is no evidence of bias, even after combining
20 lines of sight. \tom{This result is a sanity check: if we had chosen lines
of sight with identical $\kappah^{\rm med}$ rather than $\kappah^{\rm
med}\pm0.003$ randomly selected lines of sight would have zero bias by
construction}. If lines of sight hosting time delay lenses are truely random, 
our method allows $\kappax$ to be reconstructed without inducing a bias (under
the basic assumptions that the universe is like the calibration simulation,
and the correct stellar-mass to halo-mass relation has been used).

@@

We find that selecting the most tightly constrained lines of sight in this way
does not introduce a significant bias:  our reconstruction is most accurate
for the lowest $\kappax$ lines of sight. Combining 20 such systems, the bias
is at the 0.001 level.

@@

\phil{We can test the impact of such a shear selection by selecting only lines
of sight with an external shear above some threshold.  We define a somewhat
extreme selection  of $\gammax>0.05$, and find that in this sample, $\kappax$
would be systematically underestimated at the $\sim$0.008 level (corresponding
to a 0.8\% systematic error in distance).}  In the absence of other sources of
uncertainty, this systematic error would be significant in a sample of just 5
lenses;  approximating strong-lens modelling uncertainties as an unbiased
Gaussian with width 4\% on time-delay distance a 0.8\% bias would be
significant for $\sim$20 lenses. \flag{We said 3\% earlier! Can you check for
consistency please? And we only need to explain the Gaussian convolution
once.} This systematic error is mitigated if only the lines of sight with both
$\gammax>0.05$ and well constrained $\pr(\kappax|\data)$ are used, but in this
case a bias at the $\sim$0.005 level remains.

@@

% \comments{
% We have shown that reconstructing the matter due to halos along any line of
% sight can improve the constraints on the external convergence along that line
% of sight, with, depending on sample selection,  the width of
% $\pr(\kappax|\data)$ being up to a factor of two smaller than the global
% $\pr(\kappax)$, and  we have quantified the residual bias in the $\kappax$
% estimates under our calibration assumptions. We now revisit these assumptions.
% }

@@

Clusters, filaments and voids are more difficult to model, but simple forms
could in principle be derived from the structures found in the simulations,
and included in the halo model in some way.  Most importantly for the highest
$\kappax$ lines of sight will be a better group and cluster model:
\citet{MomchevaEtal2006} and  \citet{WongEtal2011} recommend using models
where dark mass is assigned to both galaxy-scale halos and halos for the
groups and clusters they occupy: to make this work, one could incorporate a
group and cluster finder, and then infer halo mass from optical richness
\citep[\eg][]{MaxBCG} in tandem with BCG stellar mass. \tom{Meanwhile,
improving the modelling of voids could be done using suggestions by
\citet{voids2} or \citet{voids1} combined with a void finder such as in
\citet{voids3}. Combining an advanced halo model with a void model will allow
a direct test of the halo approximation: using real photometric data one could
reconstruct lines of sight and compare against measured weak lensing shears.
This would ensure that the reconstruction model reproduces the contents of the
real universe, rather than a particular simulation.}

% \comments{Interestingly, we have found that it is the most empty lines of sight that 
% can be reconstructed with the highest precision. $\kappax$ estimates for empty
% lines of sight receive few contributions from halos, and a large contribution
% from voids; since they have the tightest PDFs after the reconstruction it
% seems that our model's biggest uncertainties are driven by naively
% reconstructing halos rather than neglecting voids. With a photometric
% reconstruction of the field there is a significant broadening, but the
% resulting precision is still typically 50\%  higher than using the global
% $\kappax$ distribution.
% 
% Since the uncertainty on $\pr(\kappax|\data)$ is $\sim$0.01 even
% with perfect knowledge of halo mass and redshift, it seems that modelling uncertainties
% from voids, deviations from sphericity and dark matter clumping within the
% main halo dominate the total $\kappax$ uncertainty even with very
% uncertain stellar masses. Inferring halo ellipticity and dark matter
% clumping will likely always remain a difficulty for line of sight
% reconstruction; but as \Fref{fig:magcut} shows, observing deeper
% than 24th magnitude does not significantly help the reconstruction.
% }

% \comments{For a small fraction of lines of sight, $\pr(\kappax|\data)$ remains very
% broad even after applying our reconstruction; these lines of sight are
% typically the most over-dense lines of sight in the universe. For
% time-delay cosmography the most observationally expensive tasks are the
% light-curve monitoring and high resolution follow-up imaging and spectroscopy,
% while making photometric observations of a
% 4$\times$4 arcminute patch of sky survey down to 24th magnitude is
% relatively cheap. We have shown that with a single epoch observation of
% the region around a strong lens it is possible to infer $\pr(\kappax|\data)$.
% If a line of sight has a broad $\pr(\kappax|\data)$ it can be rejected {\it
% before} the investment of follow-up. We find that
% selecting only the lines of sight with the most tightly constrained
% $\pr(\kappax)$ does not induce a bias in $\kappax$, and hence distance.
% For the most over-dense lines of sight,
% the precision of $\pr(\kappax|\data)$ is improved noticeably by the addition
% of spectroscopic information; currently insufficient time-delay systems are
% know to choose a subset based on their reconstructed $\pr(\kappax|\data)$.
% Spectroscopic coverage will definitely be useful for extracting cosmology from
% over-dense lines of sight.
% }
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \comments{%
% \begin{figure}
% \includegraphics[width=\columnwidth]{figs/biasplots3.eps}
% \caption{Convergence estimation accuracy using the full  $\pr(\kappax|\kappah)$
% distribution, rather than its median.  As in \Fref{fig:biasplots}, solid,
% dashed and dotted lines show combinations of 20, 5 and 2 lines of sight
% respectively, while green lines show the inferences 
% from photometry alone, red lines from photometry and targeted spectroscopy 
% (assuming a spectroscopic redshift for all
% halos with $i<23$ and all halos within 1 arcminute and $i<24$), and blue lines
% from the unrealistic assumption of a spectroscopic redshift for all galaxies
% in the field. }
% \label{fig:no-median-tricks}
% \end{figure}
% }%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% \comments{ One
% significant disadvantage of these types of method is that they impose a
% bottleneck in the flow of information. Once you have compressed to one number,
% it is difficult to incorporate more information later; you have to recompute
% all the calibrations. Using the full PDF for $\kappah$ is the more correct
% thing to do: at present, however, it suffers from significant bias in the
% photometric data case. \Fref{fig:no-median-tricks} shows how $\kappax$ is
% overestimated, as a result of the high scatter in halo mass for objects with 
% high stellar mass. As has already been noted, 
% future work should focus on increasing the precision of the halo
% mass estimates. 
% }

In this work, we have used  the \MS as a calibration tool. If the real
Universe is not like the \MS, then this calibration will be incorrect.
Repeating this  study using a different simulation to make the test catalogues
would allow us to quantify the sensitivity of our results to the simulation
used. Similarly the use of a different semi-analytic model to paint galaxies
onto halos, and an alternative stellar mass--halo mass relation would both
enable similar exploration of these systematic errors. \citet{SuyuEtal2010}
used the relative over-density to mitigate this effect, but it does not
completely remove the dependence on simulation. Future work should include
investigating the robustness of the reconstruction to these systematic errors.

% \comments{In particular, the method presented here is tied to the simulations at two
% points: the prior PDF for $\kappah$, and the conditional distribution
% $\pr(\kappax|\kappah)$ used to convert halo model convergence to total
% convergence. The first tie could potentially be avoided if the halo mass
% estimation gave rise to a reasonable $\pr(\kappah|\data)$ on its own.
% Possibilities for improving the precision of the halo mass estimation are
% discussed above. The second tie is more difficult, since breaking it means
% modelling the unseen mass structures and voids. It
% should be possible to infer hierarchically  the parameters of a flexible model
% for these things from the data itself, as demonstrated by
% \citet{KarpenkaEtal2012} (albeit under the assumption of scatter-free scaling
% relations) -- how this affects the cosmographic precision is to be
% investigated. When working with wide field imaging surveys,  weak lensing shears will be
% available around every line of sight:  fitting these with the halo models
% described here in order to constrain the hyper-parameters of the model presents a 
% possible test of our halo model prescription against real observations.
% }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\item Selecting from the highest precision 33\% of the reconstructed lines of
sight results in a sample with low bias, $\sim$0.1\% in terms of the
time-delay distance. \tom{For these lines of sight the reconstructed
$\pr(\kappax|\data)$ induces a statistical uncertainty of $\sim$1\% on
individual time-delay distances.}

### 2012-12-04 doc/references.tex
acted on phils comments
response to Tom's edits

@@

@@

@@

>>>>

\bibitem[Foster 
\& Nelson(2009)]{voids3} Foster, C., \& Nelson, L.~A.\ 2009, \apj, 699, 1252

@@

\bibitem[Higuchi et al.(2012)]{voids1} Higuchi, Y., Oguri, M., 
\& Hamana, T.\ 2012, arXiv:1211.5966

@@

\bibitem[Krause et al.(2012)]{voids2} Krause, E., Chang, 
T.-C., Dor{\'e}, O., \& Umetsu, K.\ 2012, arXiv:1210.2446

@@

\bibitem[\protect\citeauthoryear{Wyithe et al.}{2011}]{WyitheEtal2011} Wyithe, J.~S.~B., Yan, 
H., Windhorst, R.~A., \& Mao, S.\ 2011, \nat, 469, 181
